### 2022-06-10 paper.tex
fixed two citations

To demonstrate an improvement of results, we compare previously published analyses described above with our analysis method.
First, we use a dataset by \citealp{Argelaguet_2019}. In their work, they generated a single cell methylome dataset of mouse embryo cells across different stages of development (E4.5 – E7.5). They calculated the correlation between RNA expression intensity and corresponding DNA methylation at the promoter region to elucidate the implications of the epigenome on RNA expression during development. As promoters, they used DNA regions of $\pm$ 2kb around transcriptions start sites (TSS).

@@

To further demonstrate the benefit of detecting variable methylation regions, we used a second dataset, published by \citealp{Luo_2017}. In their work, \citealp{Luo_2017} generated methylomes from single cell neuronal nuclei in mice to identify genomic regulatory elements across neuron types. To compare how well cell types are separated by different analysis techniques, we processed the same dataset into 2-dimensional UMAPs several times. Subsequently, we compared the results by assigning a score to each output based on cell type separation in the final 15 principal components. We see an improvement of cell separation when comparing the analysis of variable methylation regions to 100 kb regions (\ref{figure:score}A and C). The difference of using shrunken residuals instead of methylation fractions to assess the methylation per region is not as significant. However, when emulating a sparse dataset with less cells by randomly reducing the dataset to 500 cells, it becomes clear that using shrunken residuals over methylation fractions also improves the outcome of our analysis (\ref{figure:score}B and D).

\begin{figure}[!htp]
	A\\
	\hspace{.3cm}\includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/complete_135k_12cm_log.pdf} \\
	B\\
	\hspace{.3cm}\includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/cell500_20k_12cm_log.pdf} \\
	C\\
	\hspace{.39cm}\includegraphics[width=\columnwidth]{part_leonie_git/leonie_plots/UMAP_fulldataset.pdf}
	D\\
	\hspace{.39cm}\includegraphics[width=\columnwidth]{part_leonie_git/leonie_plots/UMAP_reduceddataset.pdf}
	\caption{\small \textbf{Analysis comparison}. All plots are based on the data by \citealp{Luo_2017}. \textbf{(A)} Nearest neighbor count score for complete dataset using four different analysis techniques and a variable number of regions as input. Exemplary UMAPs for encircled scores are shown in (C). \textbf{(B)} Nearest neighbor count score for the same dataset reduced to 500 cells using four different analysis techniques and a variable number of regions as input. Exemplary UMAPs for encircled scores are shown in (D). \textbf{(C)} Two exemplary UMAPs using 1000 regions of the full dataset and methylation fraction of 100 kb windows (left) or shrunken residuals of variable methylation regions (right) \textbf{(D)} Two exemplary UMAPs using 1000 regions of the reduced dataset and methylation fraction of 100 kb windows (left) or shrunken residuals of variable methylation regions (right).}
	\label{figure:score}
\end{figure}

>>>>

To demonstrate an improvement of results, we compare previously published analyses described above with our analysis method.
First, we use a dataset by \citet{Argelaguet_2019}. In their work, they generated a single cell methylome dataset of mouse embryo cells across different stages of development (E4.5 – E7.5). They calculated the correlation between RNA expression intensity and corresponding DNA methylation at the promoter region to elucidate the implications of the epigenome on RNA expression during development. As promoters, they used DNA regions of $\pm$ 2kb around transcriptions start sites (TSS).

@@

To further demonstrate the benefit of detecting variable methylation regions, we used a second dataset, published by \citealp{luo2017single}. In their work, \citealp{luo2017single} generated methylomes from single cell neuronal nuclei in mice to identify genomic regulatory elements across neuron types. To compare how well cell types are separated by different analysis techniques, we processed the same dataset into 2-dimensional UMAPs several times. Subsequently, we compared the results by assigning a score to each output based on cell type separation in the final 15 principal components. We see an improvement of cell separation when comparing the analysis of variable methylation regions to 100 kb regions (\ref{figure:score}A and C). The difference of using shrunken residuals instead of methylation fractions to assess the methylation per region is not as significant. However, when emulating a sparse dataset with less cells by randomly reducing the dataset to 500 cells, it becomes clear that using shrunken residuals over methylation fractions also improves the outcome of our analysis (\ref{figure:score}B and D).

\begin{figure}[!htp]
	A\\
	\hspace{.3cm}\includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/complete_135k_12cm_log.pdf} \\
	B\\
	\hspace{.3cm}\includegraphics[width=.7\columnwidth]{part_leonie_git/leonie_plots/cell500_20k_12cm_log.pdf} \\
	C\\
	\hspace{.39cm}\includegraphics[width=\columnwidth]{part_leonie_git/leonie_plots/UMAP_fulldataset.pdf}
	D\\
	\hspace{.39cm}\includegraphics[width=\columnwidth]{part_leonie_git/leonie_plots/UMAP_reduceddataset.pdf}
	\caption{\small \textbf{Analysis comparison}. All plots are based on the data by \citealp{luo2017single}. \textbf{(A)} Nearest neighbor count score for complete dataset using four different analysis techniques and a variable number of regions as input. Exemplary UMAPs for encircled scores are shown in (C). \textbf{(B)} Nearest neighbor count score for the same dataset reduced to 500 cells using four different analysis techniques and a variable number of regions as input. Exemplary UMAPs for encircled scores are shown in (D). \textbf{(C)} Two exemplary UMAPs using 1000 regions of the full dataset and methylation fraction of 100 kb windows (left) or shrunken residuals of variable methylation regions (right) \textbf{(D)} Two exemplary UMAPs using 1000 regions of the reduced dataset and methylation fraction of 100 kb windows (left) or shrunken residuals of variable methylation regions (right).}
	\label{figure:score}
\end{figure}
